\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{hyperref}

\title{Four Target-Selector Modes for the Safe MPC Filter}
\author{Repository Implementation Note}
\date{\today}

\begin{document}
\maketitle

\section{Overview}
The safe MPC filter in this repository uses an offset-free augmented model
\begin{equation}
z_k = \begin{bmatrix} x_k \\ d_k \end{bmatrix},
\qquad
z_{k+1} =
\begin{bmatrix}
A & B_d \\
0 & I
\end{bmatrix}
z_k +
\begin{bmatrix}
B \\ 0
\end{bmatrix} u_k,
\qquad
y_k = Cx_k + C_d d_k,
\end{equation}
with the current augmented estimate $\hat z_k = [\hat x_k;\hat d_k]$. The target-selector layer computes a steady package
\[
(x_s,u_s,d_s,y_s,y_{c,s},r_s)
\]
used by the downstream MPC objective, Lyapunov center, and safety-correction QCQP. The present implementation supports four selector modes behind one common entrypoint
\[
\texttt{prepare\_filter\_target(..., selector\_mode, config, prev\_target, H)}.
\]

\section{Shared Notation}
Let $H$ select controlled outputs so that
\[
y_{c,s} = Hy_s, \qquad r_s \in \mathbb{R}^{n_c}.
\]
The tightened input and output bounds are
\[
u_{\ell}^{t} = u_{\ell}+u_{\text{tight}}, \qquad u_{h}^{t} = u_h-u_{\text{tight}},
\]
\[
y_{\ell}^{t} = y_{\ell}+y_{\text{tight}}, \qquad y_{h}^{t} = y_h-y_{\text{tight}}.
\]
The steady-state dynamics are written as
\[
(I-A)x_s - Bu_s - B_d d_s = 0.
\]
Soft output bounds are implemented with nonnegative slack variables and quadratic penalties. Warm start uses the previous accepted target $(x_s^{\text{prev}},u_s^{\text{prev}},d_s^{\text{prev}})$ when available.

\section{Mode 0: Current Exact/Fallback Frozen-$d$}
Mode 0 preserves the existing two-stage Rawlings-style selector. The disturbance is frozen at
\[
d_s = \hat d_k.
\]
Stage 1 solves
\begin{align}
\min_{x_s,u_s}\quad &
(u_s-u_{\text{nom}})^\top R_u (u_s-u_{\text{nom}})
 + x_s^\top Q_x x_s + \Phi_y^{\text{soft}}(y_s) \\
\text{s.t.}\quad &
(I-A)x_s - Bu_s - B_d\hat d_k = 0, \\
& y_s = Cx_s + C_d \hat d_k, \\
& y_{c,s} = H y_s = y_{sp}, \\
& u_{\ell}^{t} \le u_s \le u_{h}^{t},
\end{align}
plus optional output bounds. If Stage 1 is rejected, Stage 2 solves
\begin{align}
\min_{x_s,u_s}\quad &
(y_{c,s}-y_{sp})^\top T_y (y_{c,s}-y_{sp})
 + (u_s-u_{\text{nom}})^\top R_u (u_s-u_{\text{nom}}) \\
& + x_s^\top Q_x x_s
 + (x_s-x_s^{\text{prev}})^\top Q_{\Delta x}(x_s-x_s^{\text{prev}})
 + (u_s-u_s^{\text{prev}})^\top R_{\Delta u}(u_s-u_s^{\text{prev}}) \nonumber \\
& + \Phi_y^{\text{soft}}(y_s)
\end{align}
under the same steady-state and bound constraints. In this mode
\[
r_s = y_{c,s}.
\]

\section{Mode 1: Free-Disturbance Prior}
Mode 1 treats the pseudo steady disturbance as an optimization variable, not a frozen truth:
\[
d_s \text{ is optimized and penalized around } \hat d_k.
\]
The single-stage problem is
\begin{align}
\min_{x_s,u_s,d_s}\quad &
\|r_s-y_{sp}\|_{Q_r}^2
\!+\!\|u_s-u_{\text{nom}}\|_{R_u}^2
\!+\!\|x_s-x_s^{\text{prev}}\|_{Q_{\Delta x}}^2 \\
&+\|u_s-u_s^{\text{prev}}\|_{R_{\Delta u}}^2
\!+\!\|d_s-\hat d_k\|_{Q_d}^2
\!+\!\Phi_y^{\text{soft}}(y_s)
\end{align}
subject to
\begin{align}
(I-A)x_s - Bu_s - B_d d_s &= 0, \\
y_s &= Cx_s + C_d d_s, \\
y_{c,s} &= H y_s, \\
r_s &= y_{c,s}, \\
u_{\ell}^{t} \le u_s &\le u_{h}^{t}.
\end{align}
Optionally,
\[
\|d_s-\hat d_k\|_{\infty} \le \Delta_d.
\]
This mode is designed for the case where the disturbance estimate is a good bias prior but not a trustworthy steady-state coordinate.

\section{Mode 2: Compromised Reference}
Mode 2 keeps the disturbance frozen but explicitly promotes the admissible steady controlled output to the official target. The problem is
\begin{align}
\min_{x_s,u_s}\quad &
\|r_s-y_{sp}\|_{Q_r}^2
\!+\!\|u_s-u_{\text{nom}}\|_{R_u}^2
\!+\!\|x_s-x_s^{\text{prev}}\|_{Q_{\Delta x}}^2 \\
&+\|u_s-u_s^{\text{prev}}\|_{R_{\Delta u}}^2
\!+\!\Phi_y^{\text{soft}}(y_s)
\end{align}
subject to
\begin{align}
(I-A)x_s - Bu_s - B_d \hat d_k &= 0, \\
y_s &= Cx_s + C_d \hat d_k, \\
y_{c,s} &= H y_s, \\
r_s &= y_{c,s}, \\
u_{\ell}^{t} \le u_s &\le u_{h}^{t}.
\end{align}
The downstream rule in this repository is: when this mode is selected, the upstream MPC candidate, the QCQP output term, and the backup MPC all use $r_s$ as their tracking target whenever dimensions are compatible.

\section{Mode 3: Single-Stage Robust SSTP}
Mode 3 removes the exact/fallback switch and always solves one softened steady-state target problem. In the present implementation, $d_s$ is frozen at $\hat d_k$ and both the steady-state equation and the controlled-output reference are softened:
\begin{align}
\min_{x_s,u_s,\epsilon_x,\epsilon_y}\quad &
\|r_s-y_{sp}\|_{Q_r}^2
\!+\!\|u_s-u_{\text{nom}}\|_{R_u}^2
\!+\!\|x_s-x_s^{\text{prev}}\|_{Q_{\Delta x}}^2 \\
&+\|u_s-u_s^{\text{prev}}\|_{R_{\Delta u}}^2
\!+\!\rho_x \|\epsilon_x\|_1
\!+\!\rho_y \|\epsilon_y\|_1
\!+\!\Phi_y^{\text{soft}}(y_s)
\end{align}
subject to
\begin{align}
(I-A)x_s - Bu_s - B_d \hat d_k &= \epsilon_x, \\
y_s &= Cx_s + C_d \hat d_k, \\
y_{c,s} &= H y_s, \\
r_s &= y_{c,s} + \epsilon_y, \\
u_{\ell}^{t} \le u_s &\le u_{h}^{t}.
\end{align}
Large $\rho_x$ and $\rho_y$ keep the problem close to a true steady-state target calculation while avoiding the binary exact/fallback switch.

\section{Downstream Consistency Rules}
The safe filter now uses the selector result through an \emph{effective target}:
\begin{itemize}
\item current target if the current solve succeeded;
\item otherwise the last valid target if the backup policy is \texttt{last\_valid};
\item otherwise no target.
\end{itemize}
The tracking target used by the upstream MPC and QCQP is mode-consistent:
\begin{center}
\begin{tabular}{lll}
\toprule
Mode & Default downstream target & Disturbance treatment \\
\midrule
Mode 0 & current policy (\texttt{raw} or admissible) & frozen at $\hat d_k$ \\
Mode 1 & selector reference if available & optimized with prior \\
Mode 2 & forced selector reference $r_s$ & frozen at $\hat d_k$ \\
Mode 3 & selector reference $r_s$ by default & frozen at $\hat d_k$ \\
\bottomrule
\end{tabular}
\end{center}

\section{Optimization Methods}
All four modes are convex and are solved with CVXPY through a common solver-preference sequence. The code resets variable values, applies initial values from the previous accepted target when available, and tries the configured solver list in order. Accepted statuses are currently
\[
\{\texttt{optimal},\texttt{optimal\_inaccurate}\}.
\]
Mode 0 uses an explicit two-stage solve. Modes 1--3 use a single-stage convex problem.

\section{Implementation Mapping}
The code mapping is:
\begin{itemize}
\item \texttt{Lyapunov/target\_selector.py}
  \begin{itemize}
  \item \texttt{prepare\_filter\_target(...)}: common API and mode dispatch
  \item \texttt{build\_target\_selector\_config(...)}: defaults and override merge
  \item \texttt{compute\_ss\_target\_refined\_rawlings(...)}: preserved Mode 0 core
  \item \texttt{\_compute\_target\_mode2\_compromised\_reference(...)}
  \item \texttt{\_compute\_target\_mode1\_free\_disturbance\_prior(...)}
  \item \texttt{\_compute\_target\_mode3\_single\_stage\_robust\_sstp(...)}
  \end{itemize}
\item \texttt{Simulation/run\_mpc\_lyapunov.py} and \texttt{Simulation/run\_rl\_lyapunov.py}
  \begin{itemize}
  \item build selector config once per rollout
  \item select the downstream tracking target consistently by mode
  \item propagate selector diagnostics into the debug stream
  \end{itemize}
\item \texttt{Lyapunov/safety\_debug.py}
  \begin{itemize}
  \item stores \texttt{selector\_mode}, \texttt{r\_s}, the disturbance-deviation metric $d_s-\hat d$, and the decomposition
  \item saves plots for $C x_s$, $C_d d_s$, and $y_s$
  \end{itemize}
\end{itemize}

\section{Advantages, Risks, and Expected Failure Modes}
\textbf{Mode 0} is the baseline for reproducibility, but it can spend most of its life in fallback under estimator drift. \textbf{Mode 1} is more flexible when $\hat d_k$ is not a trustworthy steady-state disturbance, but it can blur state-versus-disturbance identifiability. \textbf{Mode 2} removes target mismatch across layers, but if the selector itself is poor then the entire loop consistently tracks a poor admissible target. \textbf{Mode 3} removes explicit stage switching, but overly small $\rho_x$ or $\rho_y$ can turn the steady target into a numerically convenient but physically weak surrogate.

\section{References}
\begin{thebibliography}{9}
\bibitem{rawlings}
J. B. Rawlings, D. Q. Mayne, and M. Diehl,
\emph{Model Predictive Control: Theory, Computation, and Design},
2nd ed., Nob Hill Publishing, 2020.

\bibitem{muske}
K. R. Muske and T. A. Badgwell,
``Disturbance modeling for offset-free linear model predictive control,''
\emph{Journal of Process Control}, 2002.

\bibitem{kuntz}
J. Kuntz,
\emph{Towards a Turnkey Model Predictive Controller},
PhD thesis, University of California, Santa Barbara, 2024.

\bibitem{yan}
X. Yan et al.,
``A predictive control framework with guaranteed safety,''
\emph{Automatica}, 2023.

\bibitem{kassmann}
D. Kassmann, T. A. Badgwell, and R. B. Hawkins,
``Robust steady-state target calculation for model predictive control,''
\emph{AIChE Journal}, 2000.
\end{thebibliography}

\end{document}
